My title

### 2023-10-03 outputs/paper/paper.qmd
Update the YAML for each chunk

## Load packages
library(tidyverse)
library(palmerpenguins)
library(here)
library(kableExtra)

@@

# Data {#sec-data}
The data will be used in this paper was obtained from Open Data Toronto Portal by the library `opendatatoronto`. The name of the data file is `Bicycle Thefts`. The data contains above 30,000 rows and about 30 columns. The raw data was cleaned using the open source statistical programming language with packages 1, 2, 3. In the cleaned data, it contains 11 variables which can be divided into three main categories.

## The Time of Bicycle Thefts Occurrence
The cleaned data contains very detailed information regarding the occurrence time of each theft. The smallest unit is the occurrence hour of each theft. The table below shows the summary table for the occurrence of bicycle thefts

```{r}
cleaned_data %>% 
  group_by(occ_year) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  rename(`Occurrence Year` = occ_year,
         `Count` = n) %>% 
  kable()
```

## The Places Where the Bike was stolen
The second main category in the cleaned data is describing the places of bicycle thefts. There are six main premises type of bicycle thefts, which are: Commercial, Apartment, Other, House, Educational, Transit and Other. The premises type here represent the places where the bicycle was stolen.

```{r}
cleaned_data %>% 
  group_by(premises_type) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  rename(`Premise Type` = premises_type,
         `Count` = n) %>% 
  kable()
```

## The Stolen Bicycles' Own Features
The last category of variables in the cleaned data is the features of each stolen bicycle. The features of each stolen bike include its brand, type, color and cost.

```{r}
cleaned_data %>% 
  group_by(bike_type) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  rename(`Bike Type` = bike_type,
         `Count` = n) %>% 
  kable()
```

```{r}
cleaned_data %>% 
  group_by(bike_colour) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  rename(`Bike Colour` = bike_colour,
         `Count` = n) %>% 
  kable()
```

# Results
## Compare the bicycle thefts in the recent six years
```{r, echo = FALSE}
summarized_data <- cleaned_data %>%
  filter(occ_year %in% 2017:2022) %>% 
  group_by(occ_year, occ_month) %>%
  summarize(Total_Thefts = n())

summarized_data %>% 
    mutate(occ_month = factor(occ_month, levels = month.name)) %>% 
    ggplot(aes(x = occ_month, y = Total_Thefts, group = occ_year, color =  as.factor(occ_year))) +
    geom_line(size = 1) +
    scale_colour_manual(name = "Years", 
                      values = c("grey", "black", "green", "red", "blue","skyblue")) +
    labs(title = "The number of bicycle thefts from 2017 to 2022 by month",
       x = "Month",
       y = "Number of bicycle thefts") +
    theme_minimal() +
    guides(x = guide_axis(angle = 90))
```

Descriptions here

## Compare the premises types and probability to be stolen during summer.
```{r}
cleaned_data <- cleaned_data %>% 
  mutate(stolen_summer = case_when(occ_month %in% c("May", "June", "July", "August") ~ "Yes",
                                   TRUE ~ "No"))
cleaned_data |>
  filter(occ_year >= 2020 & premises_type != "Other") %>% 
  ggplot(mapping = aes(x = premises_type, fill=stolen_summer)) +
  geom_bar() +
  theme_minimal() +
  labs(
    x = "Premises type",
    y = "Number of thefts",
    fill = "Stolen during summer?") +
  facet_wrap(vars(occ_year)) +
  guides(x = guide_axis(angle = 90)) +
  theme(legend.position = "bottom")
```

Descriptions here

@@

```{r}
mean_price <- mean(cleaned_data$bike_cost, na.rm = TRUE)
top_colors <- cleaned_data %>%
  count(bike_colour) %>%
  arrange(desc(n)) %>%
  slice(1:4) %>%
  pull(bike_colour)
top_bike_types <- cleaned_data %>%
  count(bike_type) %>%
  arrange(desc(n)) %>%
  slice(1:6) %>%
  pull(bike_type)

@@

Descriptions here

@@

For the summer month, we can observe a “pendulum movement“ in bicycle thefts. The number of thefts alternates annually, increasing one year and then decreasing the next. This trend repeats consistently from 2017 to 2022. Comparing 2022 to 2021, we see that the number of bicycle thefts during summer of 2022 started to increase again comparing to 2021. We may expect that the number of bicycle thefts decrease during the past summer. In addition, we can observe that most decrease is between 2020 and 2021, where there was about 200 decrease in the thefts during summer months. However, the potential reason behind this dramatic decrease is COVID, we can not use this to justify the number changed in bicycle thefts this year.

@@

In terms of bicycle own features. Even though one third of stolen bicycles are black, we can not say that black bicycles are more likely to be lost as number of black bicycles is high. However, one important trend is that the price of a bike will not influence its likelihood to be stolen. That said, a cheap and an expensive bike has the same probability to be stolen. Don’t take chances that your bike won’t be stolen even though it is cheap.

>>>>

## Load packages
library(tidyverse)
library(palmerpenguins)
library(here)
library(kableExtra)
library(knitr)

@@

# Data {#sec-data}

The data will be used in this paper was obtained from Open Data Toronto Portal by the library `opendatatoronto`. The name of the data file is `Bicycle Thefts`. The data contains above 30,000 rows and about 30 columns. The raw data was cleaned using the open source statistical programming language with packages 1, 2, 3. In the cleaned data, it contains 11 variables which can be divided into three main categories.

## The Time of Bicycle Thefts Occurrence

The cleaned data contains very detailed information regarding the occurrence time of each theft. The smallest unit is the occurrence hour of each theft. The table below shows the summary table for the occurrence of bicycle thefts

```{r}
#| echo: false
#| message: false
#| label: tbl_theft_occ_year
#| tbl-cap: Summary of Bicycle Thefts Occurrence Year
cleaned_data %>% 
  group_by(occ_year) %>% 
  summarise(n = n()) %>% 
  arrange(desc(occ_year)) %>% 
  rename(`Occurrence Year` = occ_year,
         `Count` = n) %>% 
  kable()
```

Table 1 shows the number of bicycle thefts from 2009 to 2022. The difference between 2009 to 2013 and 2014 to 2022 indicate the missing data for latter years. From 2014 to 2022, we can see that number of bicycle thefts increased first from 2014 to 2015, then the bicycle thefts droped gradually since 2018. The reported bicycle thefts for last year was about 2400.

## The Premises Place Where the Bike was stolen

The second main category in the cleaned data is describing the places of bicycle thefts. There are six main premises type of bicycle thefts, which are: Commercial, Apartment, Other, House, Educational, Transit and Other. The premises type here represent the places where the bicycle was stolen.

```{r}
#| echo: false
#| message: false
#| label: a
#| tbl-cap: Summary of Bicycle Thefts Occurrence Year
cleaned_data %>% 
  group_by(premises_type) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  rename(`Premise Type` = premises_type,
         `Count` = n) %>% 
  kable()
```

Table 2 summarized the premises type of bicycle thefts. We can observe that premises types "Outside" and "Apartment" have more bicycle thefts than the rest places, meaning that people need to pay more attention on their bicycle if they park it near the condo or on the street.

## The Stolen Bicycles' Own Features

The last category of variables in the cleaned data is the features of each stolen bicycle. The features of each stolen bike include its brand, type, color and cost.

```{r}
#| echo: false
#| message: false
#| label: b
#| tbl-cap: Summary of Bicycle Thefts Occurrence Year
summary_stats <- data.frame(
  Mean = mean(cleaned_data$bike_cost, na.rm = TRUE),
  SD = sd(cleaned_data$bike_cost, na.rm = TRUE),
  Min = min(cleaned_data$bike_cost, na.rm = TRUE),
  Median = median(cleaned_data$bike_cost, na.rm = TRUE),
  Max = max(cleaned_data$bike_cost, na.rm = TRUE))

kable(summary_stats)

```

Table3 shows the bicycle price of stolen bicycles. The mean price of stolen bicycles is around 1000 dollars. However, it is surprising that the most expensive stolen bicycles worth 120,000 dollars.

```{r}
#| echo: false
#| message: false
#| label: c
#| tbl-cap: Summary of Bicycle Thefts Occurrence Year
cleaned_data %>% 
  group_by(bike_colour) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  rename(`Bike Colour` = bike_colour,
         `Count` = n) %>% 
  kable()
```

# Results

## Compare the bicycle thefts in the recent six years

```{r}
#| echo: false
#| message: false
#| label: d
#| fig-cap: The number of bicycle thefts from 2017 to 2022 by month
summarized_data <- cleaned_data %>%
  filter(occ_year %in% 2017:2022) %>% 
  group_by(occ_year, occ_month) %>%
  summarize(Total_Thefts = n())

summarized_data %>% 
    mutate(occ_month = factor(occ_month, levels = month.name)) %>% 
    ggplot(aes(x = occ_month, y = Total_Thefts, group = occ_year, color =  as.factor(occ_year))) +
    geom_line(size = 1) +
    scale_colour_manual(name = "Years", 
                      values = c("grey", "black", "green", "red", "blue","skyblue")) +
    labs(x = "Month",
         y = "Number of bicycle thefts") +
    theme_minimal() +
    guides(x = guide_axis(angle = 90))
```

The Figure 1 compares the monthly bicycle thefts from 2017 to 2022 with each line with different colors representing different years. It seems that the summer has significant higher bicycle thefts than winter in each year. In addition, it seems that the July is the most severe month of bicycle thefts in each year, ranged from 350 to 550 from the previous years. On the beginning and ending of each year, the number of bicycle thefts were always stable, keeping constant around 100.

## Compare the premises types and probability to be stolen during summer.

```{r}
#| echo: false
#| message: false
#| label: e
#| fig-cap: Summary of Bicycle Thefts Occurrence Year
cleaned_data <- cleaned_data %>% 
  mutate(stolen_summer = case_when(occ_month %in% c("May", "June", "July", "August") ~ "Yes",
                                   TRUE ~ "No"))
cleaned_data |>
  filter(occ_year >= 2020 & premises_type != "Other") %>% 
  ggplot(mapping = aes(x = premises_type, fill=stolen_summer)) +
  geom_bar() +
  theme_minimal() +
  labs(
    x = "Premises type",
    y = "Number of thefts",
    fill = "Stolen during summer?") +
  scale_fill_manual(values = c("No" = "lightcoral", "Yes" = "lightgrey")) +
  facet_wrap(vars(occ_year)) +
  guides(x = guide_axis(angle = 90)) +
  theme(legend.position = "bottom")
```

Figure 3 compares the premises type of bicycle thefts in the recent three years and indicate the proportion of bicycles that are stolen during the summer. There are three facet indicating three different years. Each bar in each facet represent the number of bicycle thefts of a premises type. The filled color in each bar shows the proportion of bicycles that are stolen during the summer, with light coral meaning NO and light grey meaning Yes.

@@

```{r}
#| echo: false
#| message: false
#| label: f
#| fig-cap: Summary of Bicycle Thefts Occurrence Year
mean_price <- mean(cleaned_data$bike_cost, na.rm = TRUE)
top_colors <- cleaned_data %>%
  count(bike_colour) %>%
  arrange(desc(n)) %>%
  slice(1:4) %>%
  pull(bike_colour)
top_bike_types <- cleaned_data %>%
  count(bike_type) %>%
  arrange(desc(n)) %>%
  slice(1:6) %>%
  pull(bike_type)

@@

Figure 3 compares the number of bicycle thefts for the top 4 colours and top 5 bicycle types. Each bar represent the corresponding number of bicycles of a bike type under different colors. In addition, each bar is filled by two colors orange and dark teal representing the proportion of bicycles that have a price above mean or below mean price.

@@

For the summer month, we can observe a "pendulum movement" in bicycle thefts. The number of thefts alternates annually, increasing one year and then decreasing the next. This trend repeats consistently from 2017 to 2022. Comparing 2022 to 2021, we see that the number of bicycle thefts during summer of 2022 started to increase again comparing to 2021. We may expect that the number of bicycle thefts decrease during the past summer. In addition, we can observe that most decrease is between 2020 and 2021, where there was about 200 decrease in the thefts during summer months. However, the potential reason behind this dramatic decrease is COVID, we can not use this to justify the number changed in bicycle thefts this year.

@@

In terms of bicycle own features. Even though one third of stolen bicycles are black, we can not say that black bicycles are more likely to be lost as number of black bicycles is high. However, one important trend is that the price of a bike will not influence its likelihood to be stolen. That said, a cheap and an expensive bike has the same probability to be stolen. Don't take chances that your bike won't be stolen even though it is cheap.
